---
title: "Occupation Recommendation Analysis"
author: "Yanjin Li"
date: "September 15, 2016"
output: html_document
---
## 1. Project Summary 

  Our project focused on job recommendation according to four interesting chosen factors, including arrival time for work, location, working hours and commuting means. We investigated their effects on income respectively and helped make a job-decision for your reference. Thus, our team has dug into the relations between following factors: 

* Part 2. Arrival Time vs. Occupation: We compared different occupations' arrival time for work and used bubble plots to show different occupations' median arrival time and median income together.

* Part 3. Location vs. Occupation:  We explored the difference in income and employees' number of different occupations in each states. 

* Part 4. Transportation vs. Occupation: We showed that people who do certain jobs and commute to work in a certain way earn more. 

* Part 5. Working Hours vs. Occupation: We analyzed how working hour affected different occupations' salaries.


```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# STEP 0. Package Setup and Initialization ========================================================================
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(scales)
library(d3heatmap)
library(RColorBrewer)
library(xlsx)
library(directlabels)
library(rJava)
library(reshape)

# Check which user is running the script
user <- Sys.getenv("HOME")

# SET YOUR LOCAL WORKING DIRECTORY HERE 
# pwd <- ...

if (user == "/Users/yanjin.li"){
  # Set working directory
  setwd("/Users/yanjin1993/Google Drive/Columbia University /2016 Fall /Applied Data Science /001 Project")
  # Set local saving directory
  save.path <- "/export_dataframes"
}
# Load data 
load("exported_dataframes/dat.pus",verbose = FALSE)

# * __SENARIO__: Housing unit/GQ person serial number 
# * __PUMA__: Public use microdata area code 
# * __ST__: State code 
# * __JWMNP__: Travel time to work 
# * __JWTR__: Means of transportation to work 
# * __INDP__: Industry recode for 2013 and later based on 2012 IND codes 
# * __JWAP__: Time of arrival at work - hour and minute 
# * __PINCP__: Total person's income (signed)
# * __OCCP__: Occupation 
```



## 2. Arrival Time vs. Occupation 

### 2.1 Arrival Time vs. Occupation
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# QIN YU (Arrival Time) ==========================================================================================
# ================================================================================================================
#Import data#
pop1<-dat.pus
data_d<-as.data.frame(pop1[,c(3,10,11,14)])
data_d<-na.omit(data_d)

#Divide Arrival Time as 6-12 & Others#
JWAPnames<-c("6-7","7-8","8-9","9-10","10-11","11-12","Others")
for(i in 70:81)
{data_d$JWAP[data_d$JWAP==i]<-JWAPnames[1]}
for(i in 82:93)
{data_d$JWAP[data_d$JWAP==i]<-JWAPnames[2]}
for(i in 94:105)
{data_d$JWAP[data_d$JWAP==i]<-JWAPnames[3]}
for(i in 106:117)
{data_d$JWAP[data_d$JWAP==i]<-JWAPnames[4]}
for(i in 118:129)
{data_d$JWAP[data_d$JWAP==i]<-JWAPnames[5]}
for(i in 130:141)
{data_d$JWAP[data_d$JWAP==i]<-JWAPnames[6]}
for(i in 1:69)
{data_d$JWAP[data_d$JWAP==i]<-JWAPnames[7]}
for(i in 142:285)
{data_d$JWAP[data_d$JWAP==i]<-JWAPnames[7]}

#Define Occupations#
OCP<-c("MGR","BUS","FIN","CMM","ENG","SCI","CMS","LGL","EDU","ENT","MED","HLS","PRT"
       ,"EAT","CLN","PRS","SAL","OFF","FFF","CON","EXT","RPR","PRD","TRN","MIL")
OCP<-as.vector(OCP)

#Define State names#
state<-read.csv("/Users/yanjin1993/Google Drive/Columbia University /2016 Fall /Applied Data Science /001 Project/original_dataframes/statenames.csv",header=T)
stnums<-as.vector(state[,1])
stnames<-c("AL","AK","AZ","AR","CA","CO","CT","DE","DC","FL","GA","HI","ID","IL","IN"
           ,"IA","KS","KY","LA","ME","MD","MA","MI","MN","MS","MO","MT","NE","NV","NH"
           ,"NJ","NM","NY","NC","ND","OH","OK","OR","PA","RI","SC","SD","TN","TX","UT"
           ,"VT","VA","WA","WV","WI","WY","PR")
stnames<-as.vector(stnames)
for(i in 1:52)
{data_d$ST[data_d$ST==stnums[i]]<-stnames[i]}

#Plot Arrival Time v.s. Occupation#
data_OCCP<-matrix(0,ncol=8,nrow=25)
for(i in 1:25)
{for(j in 2:8)
{data_OCCP[i,j]<-sum((data_d$occupation==OCP[i])&(data_d$JWAP==JWAPnames[j-1]))}}
data_OCCP<-data_OCCP[-c(1:3),]
colnames(data_OCCP)<-c("Occupation",JWAPnames)
data_OCCP<-as.data.frame(data_OCCP)
data_OCCP1<-melt(data_OCCP,id.vars="Occupation")
data_OCCP[,1]<-OCP[-c(1:3)]
data_OCCP1<-melt(data_OCCP,id.vars="Occupation")
colnames(data_OCCP1)[2]<-"ArrivalTime"
colnames(data_OCCP1)[3]<-"CumulativeProportion"
g<-ggplot(data=data_OCCP1,aes(x=Occupation,y=CumulativeProportion,fill=ArrivalTime))
g+geom_bar(position="fill",stat="identity")+ggtitle("Arrival Time v.s. Occupation")+scale_fill_manual(values=c(
"red","green","yellow","brown","orange","grey",'blue')) 
#ggplotly()
```

From this plot, we can see arrival time is different due to the variety of occupations. Most occupations require an arrival time between 4am and 12pm. Outliers include occupations in the EAT sector, like chefs or waiters, who begin their shifts after noon.


### 2.2 Arrival Time vs. States
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Plot Arrival Time v.s. State#
data_JWAP<-matrix(0,ncol=8,nrow=51)
for(i in 1:51)
{for(j in 2:8)
{data_JWAP[i,j]<-sum((data_d$ST==stnames[i])&(data_d$JWAP==JWAPnames[j-1]))}}
colnames(data_JWAP)<-c("State",JWAPnames)
data_JWAP<-as.data.frame(data_JWAP)
data_JWAP1<-melt(data_JWAP,id.vars="State")
data_JWAP[,1]<-stnames[1:51]
data_JWAP1<-melt(data_JWAP,id.vars="State")
colnames(data_JWAP1)[2]<-"ArrivalTime"
colnames(data_JWAP1)[3]<-"CumulativeProportion"
g<-ggplot(data=data_JWAP1,aes(x=State,y=CumulativeProportion,fill=ArrivalTime))
g+geom_bar(position="fill",stat="identity")+ggtitle("Arrival Time v.s. State")+scale_fill_manual(values=c(
"red","green","yellow","brown","orange","grey",'blue')) 
#ggplotly()
```

We can see clearly from the plot that most people in most states arrive at work between 4 and 8am, but In DC, unexpectedly, most of people arrive between 8am and 12pm. Basically, if you want to sleep more and get to work later in the day, you can choose DC as your ideal place to work.

### 2.3 Arrival Time vs. States/Occupation in Bubble Plots 
```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Reload data to make bubble plots#
pop2<-dat.pus
data_c<-as.data.frame(pop2[,c(3,10,11,14)])
data_c<-na.omit(data_c)

#Plot State's bubble plot#
median_income<-vector()
for(i in stnums)
{ median_income<-c(median_income,median(data_c$PINCP[data_c$ST==i]))}
median_income<-median_income[-52]
stnums1<-stnums[-52]
median_arrival<-vector()
for(i in 1:51)
{ median_arrival<-c(median_arrival,median(data_c$JWAP[data_c$ST==stnums1[i]]))}
ST<-stnames[1:length(stnums1)]
MedianArrivalTime<-median_arrival
MedianIncome<-median_income
data_c1<-data.frame(ST,MedianIncome,MedianArrivalTime)
plot_ly(data_c1,x=MedianArrivalTime,y=MedianIncome,text=paste("State: ", ST),
        mode="markers",color=MedianIncome,size=MedianIncome)
layout(title="Median Income v.s. Median Arrival Time")
```


This Figure is another piece of evidence that DC is an ideal place for work, since it shows that the people who work in DC have the latest arrival time and highest income. We suggest DC as a great place to find jobs.


```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Plot Occupation's bubble plot#
MedianArrivalTime<-vector()
for(i in 1:25)
{MedianArrivalTime<-c(MedianArrivalTime,median(data_c$JWAP[data_c$occupation==OCP[i]]))}
MedianArrivalTime<-MedianArrivalTime[-c(1:3)]
MedianIncome<-vector()
for(i in 1:length(OCP))
{ MedianIncome<-c(MedianIncome,median(data_c$PINCP[data_c$occupation==OCP[i]]))}
MedianIncome<-MedianIncome[-c(1:3)]
OCP1<-OCP[-c(1:3)]
data_c2<-data.frame(OCP1,MedianIncome,MedianArrivalTime)
plot_ly(data_c2,x=MedianArrivalTime,y=MedianIncome,text=paste("Occupation: ", OCP1),
        mode="markers",color=MedianIncome,size=MedianIncome)
layout(title="Median Income v.s. Median Arrival Time")
#End of code#
```

As we have mentioned, although the EAT sector has latest arrival time; however it also has a low salary level. Meanwhile, people who work in the LGL sector, like lawyers or judges, earn more but work earlier than others. So we need to make a choice, high salary or get up later?



## 4. Location vs. Occupation

```{r, echo = FALSE, warning=FALSE, message=FALSE, include=FALSE}
# YANJIN LI (Locations) ===========================================================================================
# =================================================================================================================
# STEP 1. Data Manipulation and Processing =====================================================================
ST.code.list <- list("AL", "AK", "AZ", "AR", "CA", "CO", "CT", "DE", "DC", "FL", "GA", 
                  "HI", "ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA", 
                  "MO", "MT", "NE", "NV", "NH", "NJ", "NM", "NY", "NC", "ND", "OH",
                  "OR", "PA", "RI", "SC", "SD", "TN", "TX", "UT", "VT", "VA", "WA",
                  "OK", "MS", "MI", "MN", "WV", "WI", "WY", "PR")


# Map Preparation: give state boundaries a white border
l <- list(color = toRGB("gray"), width = 2)
# Map Preparation: specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('gray'),
  backgroundcolor = toRGB('gray')
)

# HELPER FUNCTION 001: Get Data by Filters (OCCP, INDP)
GetInfobyST <- function(filter) {
  returnlist <- list()
  
  # Get sum data by locations 
  dat.sum <- dat.pus %>%
  select(ST, ST.code, OCCP, occupation.description, INDP, industry.description) %>%
  group_by(ST.code) %>%
  summarise(sum = n()) %>%
  arrange(ST.code)
  
  returnlist[["dat.sum"]] <- dat.sum
  
  if (filter == "OCCP") {
    # Field of interests by Locations 
    st.dat <- dat.pus %>% 
      select(ST, ST.code, OCCP, occupation.description, INDP, industry.description) %>% 
      group_by(ST.code, OCCP, occupation.description) %>%
      summarise(count = n()) %>%
      left_join(dat.sum) %>%
      ungroup() %>% group_by(ST.code) %>%
      arrange(desc(count)) %>%
      mutate(pctg = count/sum, P = percent(pctg))
  } else if (filter == "INDP") {
    st.dat <- dat.pus %>% 
      select(ST, ST.code, OCCP, occupation.description, INDP, industry.description) %>% 
      group_by(ST.code, INDP, industry.description) %>%
      summarise(count = n()) %>%
      left_join(dat.sum) %>%
      ungroup() %>% group_by(ST.code) %>%
      arrange(desc(count)) %>%
      mutate(pctg = count/sum, P = percent(pctg))
  }
  
  returnlist[["st.dat"]] <- st.dat
  
  # Top 6 of Interests by Locations 
  st.rank <- data.frame()
  for (code in ST.code.list) {
    dt <- st.dat %>% filter(ST.code == code)
    st.rank <- rbind(st.rank, dt[1:6, ])
  }
  
  if (filter == "OCCP") {
    st.rank <- st.rank %>% ungroup() %>% filter(!is.na(occupation.description)) %>% arrange(ST.code)
  } else if (filter == "INDP") {
    st.rank <- st.rank %>% ungroup() %>% filter(!is.na(industry.description)) %>% arrange(ST.code)
  }
  returnlist[["st.rank"]] <- st.rank
  
  return(returnlist)
}

# HELPER FUNCTION 002: Get Data Prepared for Maps 
GetTextbody <- function (inputdata1, inputdata2) {
  inputdata1 <- inputdata1 %>% mutate(textbody = NA)
  inputdata2 <- inputdata2 %>% mutate(text = NA)
  for (i in 1:nrow(inputdata1)){
    inputdata1$textbody[i] <- paste(inputdata1[i,c(3,7)], collapse =" ")
  }
  inputdata.temp <- inputdata1 %>% select(ST.code, sum, textbody) 
  for (i in 1:51) {
    inputdata2$text[i] <- paste(inputdata2$ST.code[i], "<br>", 
                                paste(unlist(inputdata.temp[(1+(i-1)*6):(i*6),3]), 
                                      collapse ="<br>"))
  }
  return(inputdata2)
}
```


```{r, echo = FALSE, warning=FALSE, message=FALSE, include=FALSE}
# STEP 2. Data Partion and Manipulation ========================================================================
# 2.1 Occupation (OCCP) by location (ST)
dat.occp <- GetInfobyST("OCCP")
mapdat.occp <- GetTextbody(dat.occp$st.rank, dat.occp$dat.sum)

# 2.2 Industry (INDP) by location (ST)
dat.indp <- GetInfobyST("INDP")
mapdat.indp <- GetTextbody(dat.indp$st.rank, dat.indp$dat.sum)

# 2.3 Average Income (PINCP) by OCCP
dat.income.occp <- dat.pus %>% 
  select(OCCP, occupation, occupation.description, ST.code, PINCP) %>%
  mutate(PINCP = ifelse(is.na(PINCP), 0, PINCP)) %>% 
  group_by(occupation) %>%
  summarise(avg.income = mean(PINCP)) %>%
  arrange(desc(avg.income))

# Average income by locations 
dat.income.st <- dat.pus %>% 
  select(ST.code, PINCP) %>% 
  mutate(PINCP = ifelse(is.na(PINCP), 0, PINCP)) %>%
  filter(PINCP > 0) %>%
  group_by(ST.code) %>%
  summarise(min.income = min(PINCP), max.income = max(PINCP), avg.income = mean(PINCP)) %>%
  left_join(dat.pus %>% select(ST.code, PINCP) %>% 
              group_by(ST.code) %>% summarise(pop = n())) 


```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# STEP 3. Data Visualizations ===================================================================================
# 3.1 Occupation (OCCP) by location (ST) 

# Choropleth Plots: Occupations by States 
plot_ly(mapdat.occp, z = sum, text = text, locations = ST.code, type = 'choropleth',
        locationmode = 'USA-states', color = sum, colors = "Blues",
        marker = list(line = l), colorbar = list(title = "Millions USD"), showscale = FALSE) %>%
  layout(title = '2014 US Occupation Overview by States (%)', 
         geo = g) 
```

### 3.1 Occupation Overview by US States 

The following table shows the top and lowest 10 states with the highest total number of people taking the survey per state, and their top three frequently mentioned occupations per state. 

> #### Top 5 populated states (OCCP): 

State |Top #1   |Top #2   |Top #3   |
------|---------|---------|---------|
CA| <span style="color:pink">MGR-Miscellaneous Managers</span> 1.7%| <span style="color:blue">SAL-Retail Salesperson </span> 1.5%| <span style="color:green">SAL-Cashiers</span> 1.3%|
TX| <span style="color:red">EDU-Elementry and Middle-school Teachers</span> 1.9%| <span style="color:pink">MGR-Miscellaneous Managers</span> 1.6%| <span style="color:green">SAL-Cashiers</span> 1.5%|
NY| <span style="color:red">EDU-Elementry and Middle-school Teachers</span> 1.7%| <span style="color:cyan">OFF-Secretaries and Adm Assitants </span> 1.7%| <span style="color:blue">SAL-Retail Salesperson </span> 1.5%|
FL| <span style="color:blue">SAL-Retail Salesperson </span> 1.7%| <span style="color:pink">MGR-Miscellaneous Managers</span> 1.7%| <span style="color:red">EDU-Elementry and Middle-school Teachers</span> 1.5%|
PA| <span style="color:cyan">OFF-Secretaries and Adm Assitants </span> 1.7%| TRN-Drivers/Sales Workers 1.6%| <span style="color:green">SAL-Cashiers</span> 1.5%|


From the above tables, we can conclude that the top-populated states sharing the majority of most frequent occupations, such as _Miscellanrous Managers_, _Sales_, _Cashiers_, _Elementary and Middle-school Teachers_, _Secretaries and Assistants_, _Drivers and Related-sales_, and all occpuations' corresponding percentages are all below 2.0%, which implies

(1) People in high-population states can have more occupation options, thus each occupation percentage is comparatively smaller(to small-scale states).

(2) The larger-population states do not have a specific state-building process focus; in other words, they focus on different fields of their states' improvement process, instead of one or two specific industries or fields. 


> #### Lowest 5 populated states (OCCP): 

State |Top #1   |Top #2   |Top #3   |
------|---------|---------|---------|
WY| TRN-Drivers/Sales Workers 2.2%| EDU-Elementry and Middle-school Teachers 2%| OFF-Secretaries and Adm Assitants 2%|
VT| EDU-Elementry and Middle-school Teachers 1.8%| MGR-Miscellaneous Managers 1.7%| SAL-Retail Salesperson 1.6%|
DC| __MGR-Miscellaneous Managers 4.2%__| __LGL-Lawyers, Judges, Magistrates and Other Judicial Workers 3.3%__| OFF-Secretaries and Adm Assitants 1.7%| 
AK| __CLN-Janitros and Building Cleaners 2%__| MGR-Miscellaneous Managers 1.9%| __FFF-Fishing and Hunting Workers 1.8%__|
ND| __MGR-Farmers, and Other Agricultural Managers 4.4%__| TRN-Drivers/Sales Workers 2.3%| __FFF-Miscellaneous Agricultural Workers 1.8%__|


From the above tables, we can conclude that most of the low-populated states have their own specific focus on states' industry/occupation fields. For example, in DC, two of the top three occupations are both related to government and law, while in ND, the top three occupations are more related to agriculture. From this fact, we can demonstrate that: 


(1) People in low-population states may have less occupation options, thus each occupation percentage is comparatively greater (to large-scale states).

(2) People in low-population states can have more focused industry and fields during the state-building process.  


### 3.2 Average Income Performance Overview by Locations 
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# 3.2 Industry (INDP) by location (ST) 
# Choropleth Plots: Industries by States 
plot_ly(dat.income.st, z = avg.income, text = paste("Average Income in ", ST.code, " = $", avg.income, 
                                                    "<br> Max Income = $", max.income, "<br> Min Income = $", 
                                                    min.income), 
        locations = ST.code, type = 'choropleth',
        locationmode = 'USA-states', color = avg.income, colors = "Reds",
        marker = list(line = l), colorbar = list(title = "Average Income: USD")) %>%
  layout(title = '2014 US Average Income by States (%)', geo = g)
```

From the above plots, we can tell that (regardless of occupation or industry)
* 1st-teir Average Income States: CT, NJ, MD, DC (all on the east coast of US, and comparatively smaller in population)
* 2nd-teir Average Income States: CA, WA, CO, VA, NY, NH (majority are on both the east and west coasts of the US, and comparatively larger in population)
* 3rd-teir Average Income States: TX, ND, IL and lower-tiers states are mostly from inland US. 


### 3.4 Overage Income Performance Overview by Occupations
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# 3.3 Average Income by Occupations
plot_ly(dat.income.occp, x = occupation, y = avg.income, 
  name = "Average income by Occupation", type = "bar") %>%
  layout(title = "Average Income by Occupations")
```

From the above plots, we can tell that law-related, engineering-related, finance-related, and business-related occupations earn the highest income.


### 3.5 Job Recommendation based on Locations

* People who do not have a specific focus on their career paths should consider living and working in some high-population states, such as CA, TX, NY, etc. 
* People with specific interests in some industries and focus in their career path, they should consider living and working in some low-population states with an industry-specific environment, such as lawyers in DC and farmers in ND, etc. 
* People pursuing higher income should consider living in the states such as CT, NJ, MD, and DC should choose occupations related to Law, Engineering, Finance or Business. 



## 4. Transportation vs. Occupation

In this section, we recommend the ideal job based on the heatmap and treemap below.Depending on your preferred means of transportation and your preferred income, the heatmap and treemap demonstrate a suitable occupation. This section also gives you some information about the income of a specific occupation and means of transportation.

### 4.1  Heatmap
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# YANXI CHEN (Transportation) ====================================================================================
# ================================================================================================================
library(treemap) ##load required packages
library(plyr)
library(highcharter)
library(viridisLite)
jwtr.description <- read.table(text="
01 .Car,truck,or van
02 .Bus or trolley bus
03 .Streetcar or trolley car(carro publico in Puerto Rico)
04 .Subway or elevated
05 .Railroad
06 .Ferryboat
07 .Taxicab
08 .Motorcycle
09 .Bicycle
10 .Walked
11 .Worked at home
12 .Other method", head=F, sep=".") ##Map the JWTR code to the description, the occupation already mapped in the data

dat_pus <- dat.pus
dt <- ddply(dat_pus, .(JWTR,occupation), summarize, inc= median(PINCP, na.rm=T)) ##Summarize the data with the median income by the combinations of JWTR and occupation
dt2 <- with(dt, tapply(inc,list(occupation,JWTR),unique)) ##reshape the dt data frame into a table with row names occupation and column names JWTR
dt2[9,c(3,4,6,7)] <- mean(dt2[9,],na.rm=T) ##replace missing values by the average income of this occupation EXT
dt2[14,7] <- mean(dt2[14,],na.rm=T) ##replace missing values by the average inecome of the MIL
dt3 <- dt2[-nrow(dt2),] ##delete the whole row of missing values since it cannot be repaired
colnames(dt3) <- jwtr.description[,2] ##replace JWTR code with JWTR actual description
d3heatmap(dt3[-18,],scale="column",colors="RdYlBu",height=800,width=500,na.rm = TRUE) ##draw the D3 heatmap with color scheme RdYlBu and also set the height and width of the heatmap
```

From the above heatmap above, we can easily find the preferred occupation for each means of transportation to work. On the heatmap, the deeper the blue color is , the higher he income it indicates. Conversely, the deeper the orange color on the grid, the lower income it indicates. For example, for Taxicab, the deepest blue grid is LGL, so the LGL is the best occupation for people who use Taxicab to work.  Another way to utilize this plot is to have an income reference about a specific combination of means of transportation and occupation. For instance, if you want to work in the sector of SCI and want to commute by Ferryboat, then you are expected to have an annual income of $60,500 .


### 4.2  Treemap

```{r, echo=FALSE, warning=FALSE, message=FALSE}

dt4<-ddply(dat_pus, .(JWTR,occupation),summarize, income=median(PINCP, na.rm=T), time=median(JWMNP, na.rm=T))
##add the median commute time for each combination of occupation and commute mean
dt4$time[dt4$JWTR==11]<-0 #set travel time to 0 for people who work at home
dt5<-na.omit(dt4) #remove missing values
dt5$JWTR<-jwtr.description[,2][match(dt5$JWTR,jwtr.description[,1])] ##replace JWTR code with JWTR description
tm<-treemap(dt5, index=c("JWTR","occupation"), vSize="income", vColor="time", type="value", palette=viridis(12)) ##draw a treemap, the map has been divided into 12 large grids. The 12 large grids represent the means of transportation. Each large grid contains some small grids which represent occupation;the size of each small grid is median income of that specific occupation, and the color represents the median travel time

#The following code is to transform static treemap to dynamic highchart treemap
hc_tm <- highchart() %>%
hc_add_series_treemap(tm, allowDrillToNode=T,
layoutAlgorithm="squarified",
name="tmdata") %>%
hc_title(text="Income Data") %>%
hc_tooltip(pointFormat = "<b>{point.name}</b>:<br>
median income: {point.value:,.0f}<br>
travel time:{point.valuecolor:,.0f}")
hc_tm
detach("package:plyr", unload=TRUE)
```

The above treemap has been divided into 12 large grids. The 12 large grids represent the means of transportation to work. Each large grid contains some small grids which represent occupations; the size of each small grid represents the size of the median income of that specific occupation. The larger the area of the grid is , the greater the median income. The color represents the median travel time; the darker the color, the less median travel time is needed. 

For example, if you want to choose bus or trolley bus to commute, you probably want to work in the CMM sector since you are likely to earn a higher income. Additionally, you can consider different occupations in the same means of transportation grid. It is simple to compare the median income and travel time for different occupation. Then you can choose the occupation based on your priorities.  Referencing the previous example, you may choose to work in the LGL although it has less income than CMM because it requires less commute time.

### 4.3 Summary

* Using the heatmap or treemap, individuals may have the information about the amount of income they expect to earn based on their preferred travel means and occupation.
* Also, you can find the job with the highest income that fits your travel preferences.
* You can also consider your priority when choosing an occupation taking into account your commute time and income preferences.


## 5. Working Hours vs. Occupation 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Mengyuan Tao (Working Hours) ===================================================================================
# ================================================================================================================
library(dplyr)
######read and manage data
workinghourdata<-dat.pus
workinghour<-workinghourdata[,c(1,3,6,8,11,14,15,16)]
workinghour<-workinghour[!is.na(workinghour$WKHP),]
workinghour<-workinghour[!is.na(workinghour$occupation),]
workinghour<-workinghour[workinghour$PINCP>=0,]
workinghour$PINCP<-workinghour$PINCP/1000
```


### 5.1 (Ungrouped) Working Hours by Occupations 

The occupation with longest average working hour is Derrick and Mining Machine worker, followed by Military and Lawyer.

On the country, people working in restaurant works least such as cooks or food servers or dishwashers.  Then comes people from other service industry, such as Fitness workers, Tour guides, Barbers, etc.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#######boxplot of working hour of different occupation, arrange by descending of average working hour
avgworkinghour<-workinghour%>%group_by(occupation)%>%summarise(avgwkhp=mean(WKHP))%>%arrange(desc(avgwkhp))
workinghour$occupation<-factor(workinghour$occupation,levels=avgworkinghour$occupation)
workinghour<-workinghour%>%arrange(occupation)
box_plot_occupation <- plot_ly(workinghour, x = WKHP, color = occupation, type = "box")
box_plot_occupation %>% layout(showlegend = FALSE)
```


### 5.2 (Ungrouped) Working Hours, Occupation, vs Income 

This is an overview of working hours, occupaion and income, we can see incomes diverse a lot with long working hours while diverse only a little bit with short working hours.


```{r, echo=FALSE, warning=FALSE, message=FALSE}
#####An overview of working hour, occupation and income
medincome<-workinghour%>%group_by(WKHP,occupation)%>%summarise(medincome=median(PINCP))
line_plot_nogroup<-ggplot(data=medincome, aes(x=WKHP, y=medincome, group = occupation, colour = occupation)) +
  geom_line(size=1)+
  ylab("Income (in thousand)")+
  xlab("")+
  geom_point( size=1.2)+
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.border = element_blank()
    ,axis.text=element_text(size=8)
    ,axis.ticks=element_blank()
    ,legend.title = element_blank()
  )+
  theme(legend.key.height=unit(0.7,"line")) +
  theme(legend.position = "bottom") +
  guides(colour=guide_legend(nrow=2))+
  ggtitle("Working hours vs Income")

line_plot_nogroup

#lingplotweb_nogroup<-plot_ly(avgincome, x = WKHP, y = avgincome, color = occupation)
#lingplotweb_nogroup

```

### 5.3 Grouped Working Hours 

This is a careful analysis by deviding working hours into 10 groups. So the income is increasing with the working hours when you do not work so much, but as the working hours exceed 60 hours per week, the trend becomes unclear. 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
########Careful analysis by deviding working hour into 10 groups
workinghourgroup<-workinghour
workinghourgroup$workhour<-with(workinghourgroup,ifelse(WKHP>=1&WKHP<=9,"0-9",
                                                        ifelse(WKHP>=10&WKHP<=19,"10-19",
                                                               ifelse(WKHP>=20&WKHP<=29,"20-29",
                                                                      ifelse(WKHP>=30&WKHP<=39,"30-39",
                                                                             ifelse(WKHP>=40&WKHP<=49,"40-49",
                                                                                    ifelse(WKHP>=50&WKHP<=59,"50-59",
                                                                                           ifelse(WKHP>=60&WKHP<=69,"60-69",
                                                                                                  ifelse(WKHP>=70&WKHP<=79,"70-79",
                                                                                                         ifelse(WKHP>=80&WKHP<=89,"80-89","90-99"))))))))))

workinghourgroup<-workinghourgroup%>%arrange(workhour)
workinghourgroup$workhour<-factor(workinghourgroup$workhour,levels=unique(workinghourgroup$workhour))

box_plot_income <- plot_ly(workinghourgroup, x = PINCP, color = workhour, type = "box")
box_plot_income

```

The trend of different occupation differs a lot. For example, for LGL-LAWYERS, AND JUDGES, MAGISTRATES, AND OTHER, it is increasing all the time, while for people work in medical area such as PHYSICIANS AND SURGEONS and for scientist, the income is going down as working hours going up, this phenomenon also appears on many other occupations.

If you prefer regular working hour as 8 hours per day or 40 per week, then the top 5 choices of positions are computer programmers or developers, Engineers,lawyers, scientists and physicians.

If you are fine with long working hours, you can try to be a lawyer, engineer, computer programmer or Derrick and Mining Machine worker.

It seems that for some physical work, people need to work for long hours to earn more money such as Derrick people or construction labor. The longer hours they work, the more they earn.

```{r, echo=FALSE, warning=FALSE, message=FALSE}

 medianincome<-workinghourgroup%>%group_by(workhour,occupation)%>%summarise(medincome=median(PINCP))
 # line_plot_group<-ggplot(data=averageincome, aes(x=workhour, y=avgincome, group = occupation, colour = occupation)) +
 #   geom_line(size=1)+
 #   ylab("Income(in thousand)")+
 #   xlab("")+
 #   geom_point( size=1.2)+
 #   theme_bw() +
 #   theme(
 #     plot.background = element_blank()
 #     ,panel.grid.major = element_blank()
 #     ,panel.grid.minor = element_blank()
 #     ,panel.border = element_blank()
 #     ,axis.text=element_text(size=8)
 #     ,axis.ticks=element_blank()
 #     ,legend.position="none"
 #     ,legend.title = element_blank()
 #   )+
 #   guides(colour=guide_legend(ncol=1))+
 #   ggtitle("Working hours vs Income")+
 #   scale_colour_discrete(guide = 'none') +
 #   scale_x_discrete(expand=c(0, 1)) +
 #   geom_dl(aes(label = occupation), method = list(dl.combine("first.points","last.points"), cex = 0.8))
 # 
 # line_plot_group
# 
 lineplotweb_group<-plot_ly(medianincome, x = workhour, y = medincome, color = occupation)
 
 lineplotweb_group

```




## 6. Conclusion 


## 7. Team Member and Responsibilities:

* Yanxi Chen: Transportation vs. Occupation 
* Yanjin Li: Data Processing, Location vs. Occupation, Reporting 
* Chang Liu: Presentation, Data Processing
* Mengyuan Tao: Data Processing, Working Hours vs. Occupation 
* Qin Yu: Arrival Time vs. Occupation 

